---
title: "PA1_template.Rmd"
author: "Ritesh Raj"
date: "25 August 2019"
output: html_document
References : report Wrirting In R By Roger Peng

---
---
C:\Users\RITESH\AppData\Local\Temp\Rtmp0ItrxP\downloaded_packages
title: "Peer Graded Assignment 1- Reproducible Research"
---

##Things to be done
1.Code for reading in the dataset and/or processing the data
2.Histogram of the total number of steps taken each day
3.Mean and median number of steps taken each day
4.Time series plot of the average number of steps taken
5.The 5-minute interval that, on average, contains the maximum number of steps
6.Code to describe and show a strategy for imputing missing data
7.Histogram of the total number of steps taken each day after missing values are imputed
8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

##Step 1 Code for reading in the dataset and/or processing the data


```{r, echo = TRUE}
setwd("D:/R Directory/R directory")
activity<-read.csv("D:/SQL ,C# Programs/Coursera DataSets/activity.csv")
```


Exploring the basics of this data
```{r}
dim(activity)
names(activity)
head(activity)
str(activity)
#total number of missing data
sum(is.na(activity$steps))/dim(activity)[[1]]
#transforming the date column into date format using lubridate
library(lubridate)
activity$date<-ymd(activity$date)
length(unique(activity$date))
```

##Step 2 Histogram of the total number of steps taken each day

```{r, echo = TRUE}
library(ggplot2)
DfQ2<-data.frame(tapply(activity$steps,activity$date,sum,na.rm=TRUE))
DfQ2$date<-rownames(DfQ2)
rownames(DfQ2)<-NULL
names(DfQ2)[[1]]<-"Total Steps"
png("plot1.png")
#Total Steps by date bar chart
ggplot(DfQ2,aes(y=DfQ2$`Total Steps`,x=DfQ2$date))+geom_bar(stat="identity") + ylab("Total Steps")+xlab("Date")+ggtitle("Total Steps by date")
dev.off()
ggplot(DfQ2,aes(y=DfQ2$`Total Steps`,x=DfQ2$date))+geom_bar(stat="identity") + ylab("Total Steps")+xlab("Date")+ggtitle("Total Steps by date")
#Histogram of total steps
qplot(DfQ2$`Total Steps`,geom="histogram",xlab="Total Steps",ylab="Counts",main="Total Steps Historgram")
png("plot1.1.png")
qplot(DfQ2$`Total Steps`,geom="histogram",xlab="Total Steps",ylab="Counts",main="Total Steps Historgram")
dev.off()

##Step 3  Mean and median number of steps taken each day

```{r, echo = TRUE}
library(dplyr)
DfQ3<-data.frame(round(tapply(activity$steps,activity$date,mean,na.rm=TRUE),2))
DfQ3$date<-rownames(DfQ3)
rownames(DfQ3)<-NULL
names(DfQ3)[[1]]<-"Mean Steps"
temp<-activity%>%select(date,steps) %>% group_by(date) %>% summarise(median(steps))
names(temp)[[2]]<-"Median Steps"
DfQ3$median<-temp$`Median Steps`
DfQ3<-DfQ3 %>% select(date,`Mean Steps`,median)
```

##Step 4 Time series plot of the average number of steps taken

```{r, echo = TRUE}
DfQ4<-DfQ3
DfQ4$date<-as.Date(DfQ4$date,format="%Y-%m-%d")
ggplot(DfQ4,aes(x=DfQ4$date,y=DfQ4$`Mean Steps`))+geom_bar(stat="identity")+scale_x_date()+ylab("Mean Steps by Every day")+xlab("Date")+ggtitle("Mean Steps by Date")
png("plot4.png")
ggplot(DfQ4,aes(x=DfQ4$date,y=DfQ4$`Mean Steps`))+geom_bar(stat="identity")+scale_x_date()+ylab("Mean Steps by Every day")+xlab("Date")+ggtitle("Mean Steps by Date")
dev.off()

##Step 5
##The 5-minute interval that, on average, contains the maximum number of steps

```{r, echo = TRUE}
#This is assuming that the words on average means averaging steps by date and interval
activity$interval<-factor(activity$interval)
TableQ5<-aggregate(data=activity,steps~date+interval,FUN="mean")
TableQ5<-aggregate(data=TableQ5,steps~interval,FUN="max")
```

##Step 6 Code to describe and show a strategy for imputing missing data

The common strategies for imputing missing data are 1. Constant value imputations 2. Regression model value imputations 3. Mean/mode value substitutions
```{r, echo = TRUE}
DfQ6<-activity
DfQ6$Missing<-is.na(DfQ6$steps)
DfQ6<-aggregate(data=DfQ6,Missing~date+interval,FUN="sum")
DfQ6A<-data.frame(tapply(DfQ6$Missing,DfQ6$date,sum))
DfQ6A$date<-rownames(DfQ6A)
rownames(DfQ6A)<-NULL
names(DfQ6A)<-c("Missing","date")
DfQ6A$date<-as.Date(DfQ6A$date,format="%Y-%m-%d")
DfQ6B<-data.frame(tapply(DfQ6$Missing,DfQ6$interval,sum))
DfQ6B$date<-rownames(DfQ6B)
rownames(DfQ6B)<-NULL
names(DfQ6B)<-c("Missing","Interval")
par(mfrow=c(1,2))
plot(y=DfQ6A$Missing,x=DfQ6A$date,main="Missing Value Distribution by Date")
plot(y=DfQ6B$Missing,x=DfQ6B$Interval,main="Missing Value Distribution by Interval")
table(activity$date)
```

we can see that for every interval, there are 8 values as NA. Date has 288 missing values In total there are 8 dates that have missing value.

```{r, echo = TRUE}
#Dates that have missing values 
library(lubridate)
DfQ6C<-as.data.frame(DfQ6A) %>% select(date,Missing) %>% arrange(desc(Missing))
DfQ6C<-DfQ6C[which(DfQ6C$Missing!=0),]
DfQ6C$Weekday<-wday(DfQ6C$date,label=TRUE)
DfQ6D<-activity
DfQ6D$weekday<-wday(DfQ6D$date,label=TRUE)
#Finding the mean of steps every monday, and every interval
DfQ6E<-aggregate(data=DfQ6D,steps~interval+weekday,FUN="mean",na.rm=TRUE)
#Merge the pre-imputation table DfQ6D table with the average table DfQ6E
DfQ6F<-merge(x=DfQ6D,y=DfQ6E,by.x=c("interval","weekday"),by.y=c("interval","weekday"),all.x=TRUE)
#Conditionally replacing the steps.x column NA value with the values from steps.y column value 
DfQ6F$Steps.Updated<-0
for (i in 1:dim(DfQ6F)[[1]]){
if(is.na(DfQ6F[i,3])){DfQ6F[i,6]=DfQ6F[i,5]}
else {DfQ6F[i,6]=DfQ6F[i,3]}
}
#Now simplify the imputed analytical data frame
DfQ6F <-DfQ6F  %>% select(date,weekday,interval,Steps.Updated)
names(DfQ6F)[[4]]<-"Steps"

## Step 7 Histogram of the total number of steps taken each day after missing values are imputed


```{r, echo = TRUE}
png("plot7.png")
qplot(DfQ6F$Steps,geom="histogram",main="Histogram Total steps taken after imputation",xlab="Steps",ylab="Count")
dev.off()
qplot(DfQ6F$Steps,geom="histogram",main="Histogram Total steps taken after imputation",xlab="Steps",ylab="Count")
```



## Step 8
Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r, echo = TRUE}
DfQ8<-DfQ6F
levels(DfQ8$weekday)<-c(1,2,3,4,5,6,7)
DfQ8$WDWE<-DfQ8$weekday %in% c(1,2,3,4,5)
DfQ8.1<-aggregate(data=DfQ8,Steps~interval+WDWE,mean,na.rm=TRUE)
DfQ8.1$WDWE<-as.factor(DfQ8.1$WDWE)
levels(DfQ8.1$WDWE)<-c("Weekend","Weekday")
png("plot8.png")
ggplot(data=DfQ8.1,aes(y=Steps,x=interval,group=1,color=WDWE))+geom_line() +scale_x_discrete(breaks = seq(0, 2500, by = 300))+ylab("Mean Steps")+xlab("Intervals")+ggtitle("Mean steps across intervals by Weekend and Weekday")
dev.off()
ggplot(data=DfQ8.1,aes(y=Steps,x=interval,group=1,color=WDWE))+geom_line() +scale_x_discrete(breaks = seq(0, 2500, by = 300))+ylab("Mean Steps")+xlab("Intervals")+ggtitle("Mean steps across intervals by Weekend and Weekday")
#Producing the panel plot
DfQ8.1$interval<-as.numeric(as.character(DfQ8.1$interval))
library(lattice)
xyplot(data=DfQ8.1,Steps~interval|WDWE, grid = TRUE, type = c("p", "smooth"), lwd = 4,panel = panel.smoothScatter)
library(hexbin)
hexbinplot(data=DfQ8.1,Steps~interval|WDWE, aspect = 1, bins=50)
png("plott8.1.png")
xyplot(data=DfQ8.1,Steps~interval|WDWE, grid = TRUE, type = c("p", "smooth"), lwd = 4,panel = panel.smoothScatter)
dev.off()
png("plot8.2.png")
hexbinplot(data=DfQ8.1,Steps~interval|WDWE, aspect = 1, bins=50)
dev.off()
```


